Predicting Remaining Useful Life of NASA Aircraft Engines

### 2025-02-27 index.qmd
2/26 updates

```{r, echo = FALSE}
#packages
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)

@@

test <- read.delim(testurl, header = TRUE, sep = '') %>%
  mutate(status = 0)
head(test, n = 5)
```

@@

head(train, n = 5)
```

The visual below shows a histogram and a box plot representing
distributions of survival times of the 260 unique engines in the
training data.

```{r, echo=FALSE}
distributions <- train %>% 
  group_by(id) %>%
  summarise(time = max(time))

Histogram <- ggplot(data = distributions,
       aes(x = time)) +
  geom_histogram(binwidth = 5,
                 fill = "grey")+
  labs(x = "Survival Time",
       y = "Frequency")+
  theme_bw()

@@

```{r}
summary(distributions$time)

sd(distributions$time)
```

@@

# Extract data
univ_results <- lapply(univ_models,
                       function(x){
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coefficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (",
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test",
                                        "p.value")
                          return(res)
                         })
as.data.frame(univ_results)
```

@@

@@

-   Explain your data preprocessing and cleaning steps.

>>>>

```{r, echo = FALSE}
#packages
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)
library(knitr)

@@

test <- read.delim(testurl, header = TRUE, sep = '') %>%
  mutate(status = 0)

kable(head(test, n = 5), format = "markdown")
```

@@

kable(head(train, n = 5), format = "markdown")
```

The visual below shows a histogram and a box plot representing
distributions of survival times of the 260 unique engines in the
training data. The survival times have a normal distribution

```{r}
distributions <- train %>% 
  group_by(id) %>%
  summarise(time = max(time))

summarymetrics <- summary(Minimum = min(time),
                          Median = median(time),
                          StandardDeviation = sd(time),
                          Maximum = max(time), 
                          distributions$time)

kable(tibble(summarymetrics), format = "markdown")

```{r, echo=FALSE}
Histogram <- ggplot(data = distributions,
       aes(x = time)) +
  geom_histogram(binwidth = 5,
                 fill = "grey")+
  labs(x = "Survival Time",
       y = "Frequency")+
  theme_bw()

@@

@@

# Extract data
univ_results <- lapply(univ_models,
                       function(x){
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coefficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (",
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test",
                                        "p.value")
                          return(res)
                         })
kable(as.data.frame(univ_results))
```

Assumptions

Proportional Hazards Assumption

The proportional hazards assumption states the ratio of hazard rates for any two individuals is constant over time regardless of the point in time of the measurement.

The proportional hazards assumption has been proven to be critical for evaluating prior to using Cox Proportional Hazards to model survival data.

@@

```{r}
cox <- coxph(Surv(time, status) ~ os1 + os3 + sm1 + sm2 + sm5,
             data = train)

summary(cox)
```
```{r}
cox.zph(cox)

plot(cox.zph(cox))
```

@@

-   Explain your data preprocessing and cleaning steps.
transformation applied to data to add event
